---
title: "INFX 573 Lab: MLE"
author: "Pierre Augustamar"
date: "November 3rd, 2016"
output: 
  tufte_handout:
    highlight: tango
---

\marginnote{\textcolor{blue}{Don't forget to list the full names of your collaborators!}}

# Collaborators: Derrick Phoebee

# \textbf{Instructions:}

Before beginning this assignment, please ensure you have access to R and/or RStudio. 

1. Download the `week6b_lab.Rmd` file from Canvas. Open `week6b_lab.Rmd` in RStudio (or your favorite editor) and supply your solutions to the assignment by editing `week6b_lab.Rmd`. 

2. Replace the "Insert Your Name Here" text in the `author:` field with your own full name.

3. Be sure to include code chucks, figures and written explanations as necessary. Any collaborators must be listed on the top of your assignment. Any figures should be clearly labeled and appropriately referenced within the text. 

5. When you have completed the assignment and have **checked** that your code both runs in the Console and knits correctly when you click `Knit`, rename the R Markdown file to `YourLastName_YourFirstName_lab6b.Rmd`, and knit it into a PDF. Submit the compiled PDF on Canvas.

In this lab, you will need access to the following R packages:

```{r Setup, message=FALSE}
# Load some helpful libraries
library(tidyverse)
```

Today we will explore maximum likelihood estimation. Your task is to run through the examples discussed by J.M. White and add code comments to all code chunks. Please add any comments or discussion of results as you go along as well.

# Doing Maximum Likelihood Estimation by Hand in R

By John Myles White on 4.21.2010

\noindent \href{http://www.johnmyleswhite.com/notebook/2010/04/21/doing-maximum-likelihood-estimation-by-hand-in-r/}{\textcolor{green}{Available Online}}

First, let's start with a toy example for which there is a closed-form analytic solution. We'll ignore that solution and use optimization functions to do the estimation. Starting with this toy example makes it easy to see how well an approximation system can be expected to perform under the best circumstances - and also where it goes wrong if you make poor programming decisions.

Suppose that you've got a sequence of values from an unknown Bernoulli variable like so:

```{r} 

p.parameter <- 0.8  # set the probability parameter to 0.8 
sequence <- rbinom(10, 1, p.parameter) # generates a binomial distribution with number of trials = 1, number of observations = 10, and probability of success = 0.8
```

Given the sequence, we want to estimate the value of the parameter, p, which is not known to us. The maximum likelihood approach says that we should select the parameter that makes the data most probable. For a Bernoulli variable, this is simply a search through the space of values for p (i.e [0, 1]) that makes the data most probable to have observed.

It's worth pointing out that the analytic solution to the maximum likelihood estimation problem is to use the sample mean. We'll therefore use `mean(sequence)` as a measure of the accuracy of our approximation algorithm.

How do we find the parameter numerically? First, we want to define a function that specifies the probability of our entire data set. We assume that each observation in the data is independently and identically distributed, so that the probability of the sequence is the product of the probabilities of each value. For the Bernoulli variables, this becomes the following function:

```{r}
# creating a likelihood function that takes two parameters: a probbility of events as sequence and a probability of success
likelihood <- function(sequence, p.parameter) 
{
  likelihood <- 1  # assign 1 to likelihood
 
  for (i in 1:length(sequence))  #loop through the sequence vector startig at 1 to the lenght of the sequence
  {
    if (sequence[i] == 1) # in case the sequence is 1 then likelihood will be equal to the likelihood times the parameter
    {
      likelihood <- likelihood * p.parameter  
    }
    else # otherwise likelihood estimation will be equal to the likelihood * the difference of 1 from the probability fo success
    {
      likelihood <- likelihood * (1 - p.parameter)
    }
  }
 
  return(likelihood) # returns the likelihood maximum estimation 
}
```

To do maximum likelihood estimation, we therefore only need to use an optimization function to maximize this function. A quick examination of the likelihood function as a function of p makes it clear that any decent optimization algorithm should be able to find the maximum:

```{r}
possible.p <- seq(0, 1, by = 0.001)   # generate a sequence from 0 to 1 by increasing values by 0.001
#line plot based on sequence values and the likelihood maximum extimation
qplot(possible.p,
      sapply(possible.p, function (p) {likelihood(sequence, p)}), # grouping the generated sequence with  the created likelihood function using a un-named function that takes a parameter p 
      geom = 'line',
      main = 'Likelihood as a Function of P',
      xlab = 'P',
      ylab = 'Likelihood')
```

For single variable cases, I find that it's easiest to use R's base function `optimize` to solve the optimization problem:

```{r}
# create a maximum likelihood estimation by calling the optimize method
mle.results <- optimize(function(p) {likelihood(sequence, p)}, # searches the interval from lower an upper bettern the interval 0 and 1
                        interval = c(0, 1),
                        maximum = TRUE) # macimum should be use
 
mle.results   # return the mle results 
```

Here I've used an anonymous function that returns the likelihood of our current data given a value of p; I've also specified that the values of p must lie in the interval [0, 1] and asked `optimize` to maximize the result, rather than minimize, which is the default behavior. Examining the output of `optimize`, we can see that the likelihood of the data set was maximized very near 0.7, the sample mean. This suggests that the optimization approximation can work. It's worth noting that the objective value is the likelihood of the data set for the specified value of p. The smallness of the objective for large problems can become a major problem. To understand why, it's worth seeing what happens as the size of the sample grows from 10 to 2500 samples:

```{r}

error.behavior <- data.frame() # create a data frame 
 
for (n in 10:2500) # loop through a set of numbers from 10 to 2500
{
  sequence <- rbinom(n, 1, p.parameter)    #  for each value in n generates random binomial distribution with number of trials = 1, and umber of observations = to the value of n, using the probability of success
 
  likelihood.results <- optimize(function(p) {likelihood(sequence, p)},
                                 interval = c(0, 1),
                                 maximum = TRUE)
 
  true.mle <- mean(sequence)
 
  likelihood.error <- true.mle - likelihood.results$maximum
 
  error.behavior <- rbind(error.behavior,
                          data.frame(N = n,
                                     Error = likelihood.error,
                                     Algorithm = 'Likelihood'))
}
```

As you can see, our approximation approach works great until our data set grows, and then it falls apart. This is exactly the opposite of what asymptotical statistical theory tells us should be happening, so it's clear that something is going very wrong. A quick examination of the results from the last pass through our loop makes clear what's wrong:

```{r}
sequence <- rbinom(2500, 1, p.parameter)
 
likelihood.results <- optimize(function(p) {likelihood(sequence, p)},
                               interval = c(0, 1),
                               maximum = TRUE)
 
likelihood.results
```

The likelihood of our data is numerically indistinguishable from 0 given the precision of my machine's floating point values. Multiplying thousands of probabilities together is simply not a viable approach without infinite precision. Thankfully, there's a very simple solution: replace all of the probabilities with their logarithms. Instead of maximizing the likelihood, we maximize the log likelihood, which involves summing rather than multiplying, and therefore stays numerically stable:

```{r}
log.likelihood <- function(sequence, p)
{
  log.likelihood <- 0
 
  for (i in 1:length(sequence))
  {
    if (sequence[i] == 1)
    {
      log.likelihood <- log.likelihood + log(p)
    }
    else
    {
      log.likelihood <- log.likelihood + log(1 - p)
    }
  }
 
  return(log.likelihood)
}
```

You can check that this problem is as easily solved numerically as the original problem by graphing the log likelihood:

```{r}
sequence <- c(0, 1, 1, 1, 1, 1, 1, 0, 1, 0)
possible.p <- seq(0, 1, by = 0.001)
qplot(possible.p,
      sapply(possible.p, function (p) {log.likelihood(sequence, p)}),
      geom = 'line',
      main = 'Log Likelihood as a Function of P',
      xlab = 'P',
      ylab = 'Log Likelihood')
```

And then you can rerun our error diagnostics using both approaches to confirm that the log likelihood approach does not suffer from the same numerical problems:

```{r}
error.behavior <- data.frame()
 
for (n in 10:2500)
{
  sequence <- rbinom(n, 1, p.parameter)
 
  likelihood.results <- optimize(function(p) {likelihood(sequence, p)},
                                 interval = c(0, 1),
                                 maximum = TRUE)
 
  log.likelihood.results <- optimize(function(p) {log.likelihood(sequence, p)},
                                     interval = c(0, 1),
                                     maximum = TRUE)
 
  true.mle <- mean(sequence)
 
  likelihood.error <- true.mle - likelihood.results$maximum
  log.likelihood.error <- true.mle - log.likelihood.results$maximum
 
  error.behavior <- rbind(error.behavior,
                          data.frame(N = n,
                                     Error = likelihood.error,
                                     Algorithm = 'Likelihood'),
                          data.frame(N = n,
                                     Error = log.likelihood.error,
                                     Algorithm = 'Log Likelihood'))
}
 
ggplot(error.behavior, aes(x = N, y = Error)) + geom_line(aes(group = Algorithm, color = Algorithm)) + labs(title = 'Long-Term Error Behavior \n of Two Numerical Approaches', xlab = 'Sample Size', ylab = 'Deviation from True MLE')
```

More generally, given any data set and any model, you can - at least in principle - solve the maximum likelihood estimation problem using numerical optimization algorithms. 